---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

## Download Data
Assignment is supposed to analyze data that is available on [Activity monitoring data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip). 
```{r}
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip","activity1.zip",method="curl")
unzip(zipfile="activity1.zip",overwrite=TRUE)
```

## Loading and Processing Data
```{r}
activity <- read.csv("activity.csv")
##summary(activity)
##str(activity)
## assign 0 to NA as NA might suppress the right data info too
activity$steps[is.na(activity$steps)] <- 0
## since we have to summarize data, it is better to use dplyr package. here is how we will load the library
library(dplyr)
```

## What is mean total number of steps taken per day?

Since we have to ignore the missing values for this part of the assignment, we will simply calculate the mean  


```{r}
StepsStats <- activity %>%
         group_by(date) %>%
         summarize(mean(steps),sum(steps),median(steps))
## Please we have calculated mean,total steps per day and median in the same step. To rename the columns properly use the following code
colnames(StepsStats) <- c("date","mean","totalSteps","median")
## Since steps are integer, lets round the mean to 0 decimals
StepsStats$mean <- round(StepsStats$mean,0)
## Q1- Calculate the total number of steps taken per day? display Total Steps per day
StepsStats [,c(1,3)]
## Q2- histogram for total steps per day. since the data is spread over 61 days of 2 months, hence we made a break=61
hist(StepsStats$totalSteps,breaks=61)
## Q3- mean and median per day
StepsStats [,c(1,2,4)]
```

##What is the average daily activity pattern?
```{r}
## preare data for 5 min interval
##activity$hrs <- as.character(activity$interval,length=5)
##activity$hrs [activity$interval < 10] <- sprintf("00:%02s",as.character(activity$hrs,length=5))
##activity$hrs [activity$interval >= 10 & activity$interval <100] <- sprintf("00:%2s",as.character(activity$hrs,length=5))

##activity$timeInterval = format(as.Date(paste(as.character(activity$date) , activity$hrs)),"%Y-%m-%d %H:%M")
##merge data 
##merge(StepsStats,activity,by="date")
plot(StepsStats$date,StepsStats$mean)
```


## Imputing missing values
out of time got stuck in timing format:)


## Are there differences in activity patterns between weekdays and weekends?
